---
title: "Appendix"
author: "Theo Noomah"
date: "December 9, 2018"
output: html_document
---
```{r}
library(tidyverse)
library(rvest)
library(stringdist)
```

## Overview
The following code pulls property tax records for Hyde Park, matches them on 
ownership and calculates statistics on the market concentration. Comments in the
code show how the code can be adapted to analyze any neighborhood in Cook
County.

## Subset PINs By Neighborhood
Download parcel data via:
https://datacatalog.cookcountyil.gov/GIS-Maps/ccgisdata-Parcels-2016/a33b-b59u
```{r}
parcels <- read_csv("IO Paper/parcels.csv")

hp_PINs1 <- 
  parcels %>%
  filter(PINA == 20, PINSA == 11) %>%  #PINA and PINSA are the first 4 digits of
  .$Name %>%                           #a PIN. These can be found for a given 
  as.list()                            #neighborhood using the map above.
                                       #Change these numbers to analyze a
hp_PINs2 <-                            #different neighborhood
  parcels %>% 
  filter(PINA == 20, PINSA == 14) %>% 
  .$Name %>% 
  as.list()

```

## Scraping Function
This function scrapes the Cook County Property Tax Records Portal, the following
link is an example showing what is on the page to collect: http://www.cookcountypropertyinfo.com/cookviewerpinresults.aspx?pin=20113020200000
```{r}
property_base <- 
  "http://www.cookcountypropertyinfo.com/cookviewerpinresults.aspx?pin="

pager <- function(pin) {
    profile <- read_html(str_c(property_base, pin))
    data.frame(address = profile %>% 
                 html_node("#ContentPlaceHolder1_PropertyInfo_propertyAddress") %>% 
                 html_text(),
               owner = profile %>% 
                 html_node("#ContentPlaceHolder1_PropertyInfo_propertyMailingName") %>% 
                 html_text(),
               owner_mail1 = profile %>% 
                 html_node("#ContentPlaceHolder1_PropertyInfo_propertyMailingAddress") %>%
                 html_text(),
               owner_mail2 = profile %>% 
                 html_node("#ContentPlaceHolder1_PropertyInfo_propertyMailingCityStateZip") %>%
                 html_text(),
               PIN = profile %>% 
                 html_node("#ContentPlaceHolder1_lblResultTitle") %>%
                 html_text(),
               lot = profile %>% 
                 html_node("#ContentPlaceHolder1_TaxYearInfo_propertyLotSize") %>%
                 html_text(),
               total_value = profile %>% 
                 html_node("#ContentPlaceHolder1_TaxYearInfo_propertyAssessedValue") %>%
                 html_text(),
               class = profile %>%
                 html_node("#ContentPlaceHolder1_TaxYearInfo_propertyClass") %>%
                 html_text()
               )
}
```

## Run Scraper on Neighborhood Parcel Lists

```{r}
hp_tax_data1 <- as.tibble(do.call(rbind, lapply(hp_PINs1, pager)))
hp_tax_data2 <- as.tibble(do.call(rbind, lapply(hp_PINs2, pager)))
```

## Merge Neighborhoods

```{r}
hp_tax_data <- 
  hp_tax_data1 %>% 
  rbind(hp_tax_data2)
```

## Coerce Tax Data into to Numeric

```{r}
#Hyde Park
hp_tax_data$PIN <- as.numeric(gsub("[^0-9]", "", hp_tax_data[["PIN"]]))
hp_tax_data$total_value <- as.numeric(gsub("[^0-9]", "", hp_tax_data[["total_value"]]))
hp_tax_data$lot <- as.numeric(gsub("[^0-9]", "", hp_tax_data[["lot"]]))
hp_tax_data$class <- as.numeric(gsub("[^0-9]", "", hp_tax_data[["class"]]))
hp_tax_data <-
  hp_tax_data %>%
  separate(owner_mail2, 
           into=c("city_state", "zip"),
           sep = -5)
hp_tax_data$zip <- as.numeric(hp_tax_data[["zip"]])
```

## Subset Parcels by Class

Class codes via: http://www.cookcountyassessor.com/assets/forms/classcode.pdf
```{r}
rental <- c(211, 212, 313:321, 390, 391, 396, 397, 399, 
            900, 901, 913:918, 959, 990, 991, 996, 997)

hp_rentals <- hp_tax_data %>% filter(class %in% rental)
hp_rentals$matches <- 1:nrow(hp_rentals)
```

## Function to Match Owners Within Rental Markets
```{r}
assigner_hp <- function(market_class){    #Because this function saves it's 
                                          #output as a .csv every time it runs,
for (rowa in 1:nrow(market_class)) {      #each neighborhood needs its own
    ownera <- market_class[rowa, "owner"] #function with a different name
    mail1a <- market_class[rowa, "owner_mail1"]
    zipa <- market_class[rowa, "zip"]
    matchesa <- market_class[rowa, "matches"]

    if(rowa == nrow(market_class)){break}
    
    for (rowb in (rowa+1):nrow(market_class)) {
    ownerb <- market_class[rowb, "owner"]
    mail1b <- market_class[rowb, "owner_mail1"]
    zipb <- market_class[rowb, "zip"]

    if(!zipa %in% 60000:70000 & !is.na(zipa) & !is.na(zipb)){
    if(zipa == zipb){
      market_class[rowb, "matches"] <- matchesa
    }
      
    }else if(stringdist(as.character(mail1a), as.character(mail1b)) < 1){
      market_class[rowb, "matches"] <- matchesa  
      
    }else if(stringdist(as.character(ownera), as.character(ownerb)) < 1) {
      market_class[rowb, "matches"] <- matchesa
    }
    }
    }
    assign("hp_rentals", market_class,.GlobalEnv)
    write.csv(market_class, "hp_rentals.csv")
}
```

## Run Function on Rentals in Each Neighborhood
The resulting table will have all of the given market's property tax records
with the algorithms best guess at which properties are owned by the same company
indicated by the column "matches." The values in the column are themselves 
arbitrary, but if two properties have the same number in matches they are almost
certainly owned by the same company.
```{r}
assigner_hp(hp_rentals)
```

## Summary Statistics

```{r}
# The portion of rental property value and square feet of lots owned by each of
# the biggest property owners, in descending order
hp_rentals %>% 
  group_by(matches) %>% 
  summarise(value_share = sum(total_value)/sum(.$total_value),
            lot_share = sum(lot)/sum(.$lot)) %>% 
  arrange(desc(value_share))

# The number of rental parcels, the HHI and the share of value owned by the top
# 1, 4, and 8 firms in a market
hp_rentals %>%
  filter(class %in% rental) %>% #Change out "rental"" with a specific class code
  group_by(matches) %>%         #or codes to analyze sub-markets
  summarise(value_share = sum(total_value)/sum(.$total_value),
            value_square = value_share*value_share, n = n()) %>% 
  arrange(desc(value_share)) %>% 
  summarise(n = sum(n),
            hhi = sum(value_square)*10000,
            top1 = head(value_share, 1),
            top4 = sum(head(value_share, 4)),
            top8 = sum(head(value_share, 8))
            )

# Land use in a neighborhood by square feet, but without condos
hp_tax_data %>% 
  group_by(class) %>% 
  summarise(lot_share = sum(lot, na.rm = TRUE)/sum(.$lot, na.rm = TRUE)) %>% 
  arrange(desc(lot_share))

# Percent of land used for a specific property type
hp_tax_data %>% 
  group_by(class) %>% 
  summarise(lot_share = sum(lot, na.rm = TRUE)/sum(.$lot, na.rm = TRUE)) %>% 
  filter(class %in% rental) %>% #"rental" can be changed for a specific class
  summarise(lot_share = sum(lot_share))
```
